Early Warning Signals and the Prosecutor's Fallacy

### 2013-02-13 inst/examples/fallacy.Rmd

# Code for Prosecutors Fallacy 
<!--

`r require(knitcitations)`
This code is written in the `R` language for statistical computing.  
Population dynamics are simulated using the `populationdynamics` package
`r citep(citation("populationdynamics"))` for exact simulations of 
discrete birth-death processes in continuous time using the Gillespie
agorithm `r citep("10.1021/j100540a008")`.  Early warning signals
of variance and autocorrelation, as well as the model-based estimate
of `r citet("10.1098/rsif.2012.0125")` are estimated using the 
`earlywarning` package `r citep(citation("earlywarning")`.  These
packages can be installed from Github using the `devtools` R package

```r
library(devtools)
install_github("populationdynamics", "cboettig")
install_github("earlywarning", "cboettig")
```

@@

-->

@@

``` {r simdatf eval=FALSE}
select_crashes <- function(n){
	T<- 5000
	n_pts <- n
	pars = c(Xo = 500, e = 0.5, a = 180, K = 1000, h = 200,
    i = 0, Da = 0, Dt = 0, p = 2)
	sn <- saddle_node_ibm(pars, times=seq(0,T, length=n_pts), reps=1000)
	d <- dim(sn$x1)
	crashed <- which(sn$x1[d[1],]==0)
	sn$x1[,crashed] 
}
````

@@

``` {r simdat}
sfInit(parallel=TRUE, cpu=12)
sfLibrary(populationdynamics)
sfExportAll()
examples <- sfLapply(1:20, function(i) select_crashes(50000))
dat <- melt(as.matrix(as.data.frame(examples, check.names=FALSE)))
names(dat) = c("time", "reps", "value")
levels(dat$reps) <- 1:length(levels(dat$reps)) # use numbers for reps
````

``` {r testing}
ggplot(subset(dat, reps %in% levels(dat$reps)[1:9])) + geom_line(aes(time, value)) + facet_wrap(~reps, scales="free")
````

@@

``` {r simdatf_null, eval=FALSE}
threshold <- 250
select_crashes <- function(n){
	T<- 5000
	n_pts <- n
	pars = c(Xo = 500, e = 0.5, a = 180, K = 1000, h = 200,
    i = 0, Da = 0, Dt = 0, p = 2)
	sn <- saddle_node_ibm(pars, times=seq(0,T, length=n_pts), reps=500)
	d <- dim(sn$x1)
	sn$x1[1:501,]
}
````

``` {r simdat_null, dependson="simdatf_null"}
sfInit(parallel=TRUE, cpu=12)
sfLibrary(populationdynamics)
sfExportAll()
examples <-  sfLapply(1:24, function(i) select_crashes(50000))
nulldat <- melt(as.matrix(as.data.frame(examples, check.names=FALSE)))
nulldat <- melt(examples)
names(nulldat) = c("time", "reps", "value")
levels(nulldat$reps) <- 1:length(levels(dat$reps)) 
````

@@

``` {r figure2, dependson="nullmelt"}
ggplot(dat) + geom_histogram(aes(value, y=..density..), binwidth=0.2, alpha=.5) +
 facet_wrap(~variable) + xlim(c(-1, 1)) + 
 geom_density(data=nulldat, aes(value), bw=0.2)
````

>>>>

# Code for Prosecutors Fallacy

@@

@@

``` {r simdatf}
threshold <- 250
select_crashes <- function(n){
	T<- 5000
	n_pts <- n
	pars = c(Xo = 500, e = 0.5, a = 180, K = 1000, h = 200,
    i = 0, Da = 0, Dt = 0, p = 2)
	sn <- saddle_node_ibm(pars, times=seq(0,T, length=n_pts), reps=1000)
	d <- dim(sn$x1)
	crashed <- which(sn$x1[d[1],]==0)
	sn$x1[,crashed] 
}
````

@@

``` {r parallel}
sfInit(parallel=FALSE)
sfLibrary(populationdynamics)
```

``` {r simdat, dependson="simdatf"}
sfExportAll()
examples <- sfLapply(1:20, function(i) select_crashes(50000))
dat <- melt(as.matrix(as.data.frame(examples, check.names=FALSE)))
names(dat) = c("time", "reps", "value")
levels(dat$reps) <- 1:length(levels(dat$reps)) # use numbers for reps
````

``` {r testing, dependson="simdat"}
ggplot(subset(dat, reps %in% levels(dat$reps)[1:9])) + geom_line(aes(time, value)) + facet_wrap(~reps, scales="free")
````

@@

``` {r simdatf_null}
threshold <- 250
select_crashes <- function(n){
	T<- 5000
	n_pts <- n
	pars = c(Xo = 500, e = 0.5, a = 180, K = 1000, h = 200,
    i = 0, Da = 0, Dt = 0, p = 2)
	sn <- saddle_node_ibm(pars, times=seq(0,T, length=n_pts), reps=500)
	d <- dim(sn$x1)
	sn$x1[1:501,]
}
````

``` {r simdat_null, dependson="simdatf_null"}
sfExportAll()
examples <-  sfLapply(1:10, function(i) select_crashes(50000))
nulldat <- melt(as.matrix(as.data.frame(examples, check.names=FALSE)))
nulldat <- melt(examples)
names(nulldat) = c("time", "reps", "value")
levels(nulldat$reps) <- 1:length(levels(dat$reps)) 
````

@@

``` {r fig2, dependson="nullmelt"}
ggplot(dat) + geom_histogram(aes(value, y=..density..), binwidth=0.3, alpha=.5) +
 facet_wrap(~variable) + xlim(c(-1, 1)) + 
 geom_density(data=nulldat, aes(value), adjust=2) + xlab("Kendall's tau") + theme_bw()

````

``` {r figure2, dev="CairoPS", fig.ext="eps", fig.width=6, fig.height=5, include=FALSE}
ggplot(dat) + geom_histogram(aes(value, y=..density..), binwidth=0.3, alpha=.5) +
 facet_wrap(~variable) + xlim(c(-1, 1)) + 
 geom_density(data=nulldat, aes(value), adjust=2) + xlab("Kendall's tau") + theme_bw()

````
